---
title: "Tidy Tuesday 2020.08.25 - Chopped"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(tidytuesdayR)
source("../../kp_themes/theme_kp.R")
theme_set(theme_classic() + theme_kp())
```

This week's [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-08-25/readme.md) data is all about the cooking competition TV show Chopped. It includes information about each episode's rating, air date, name, judges, contestants, and basket ingredients. The show is split into three rounds (appetizer, entree, and dessert) where contestants are challenged to create a dish using mystery basket ingredients. They may use any other multitude of ingredients in the Chopped kitchen, but are expected to highlight the basket ingredients. Judges review the dishes and "chop" (eliminate) one chef from the competition.  

The basket ingredients are often odd pairings - such as ground beef, wonton wrappers, cream of mushroom soup, and bananas (an actual appetizer round basket from episode 4). It made me wonder how often these ingredients appear together in real life. I used a [recipe dataset from Kaggle](https://www.kaggle.com/shuyangli94/food-com-recipes-and-user-interactions?select=RAW_recipes.csv) that includes thousands of recipes from Food.com to find out. Then, I extrapolated that information to create a **Chopped Basket Generator** where the selected ingredients are appropriately (un-)correlated.  

Some choice generated baskets include:  

- "Open your baskets! Your appetizer round ingredients are: peanut butter, parmigiano-reggiano cheese, apricots in juice, and fresh oregano."  
- "Open your baskets! Your entree round ingredients are: turkey, cool whip topping, ravioli, and apple pie filling."  
- "Open your baskets! Your dessert round ingredients are: buttermilk, frozen hash brown potatoes, oyster sauce, and molasses."  

  
--- 

First, let's load in the data and convert it from a string list to a long format, where each ingredient and series episode number has its own row. At the same time, let's create a dataframe where each basket's ingredients are paired together to allow for easier comparison down the road. I did this by double unnesting the same column, which feels pretty hacky, but accomplished what I needed.  

```{r chopped-data}
chopped <- tidytuesdayR::tt_load('2020-08-25')$chopped

chopped %>% select(series_episode, appetizer) %>% head(1)

chopped_lists <- chopped %>%
  select(series_episode, appetizer, entree, dessert) %>%
  mutate(appetizer = sapply(appetizer, function(x) strsplit(x, ', ')[[1]]),
         entree = sapply(entree, function(x) strsplit(x, ', ')[[1]]),
         dessert = sapply(dessert, function(x) strsplit(x, ', ')[[1]]))

chopped_lists %>% head(1)

chopped_ingredients_paired <- NULL

for(x in c('appetizer', 'entree', 'dessert')) {
  
  temp <- chopped_lists %>%
    select(series_episode, !!sym(x)) %>%
    mutate(round = x,
           ingredient = !!sym(x),
           ingredient2 = !!sym(x)) %>%
    unnest(ingredient) %>%
    unnest(ingredient2) %>%
    select(-!!sym(x)) %>%
    filter(ingredient != ingredient2) %>%
    mutate(ingredient = tolower(gsub("\\s*\\([^\\)]+\\)","",ingredient)),
           ingredient2 = tolower(gsub("\\s*\\([^\\)]+\\)","",ingredient2)))
  
  chopped_ingredients_paired <- chopped_ingredients_paired %>%
    bind_rows(temp)
  
}

chopped_ingredients_episodes <- chopped_ingredients_paired %>%
  select(-ingredient2) %>%
  unique()

chopped_ingredients <- chopped_ingredients_episodes %>%
  count(ingredient, name = 'chopped_count') 

chopped_ingredients_episodes %>%
  filter(series_episode == 1, round == 'appetizer')

chopped_ingredients_paired %>% 
  filter(series_episode == 1, round == 'appetizer')

```

Next, read in the first 10,000 rows from the recipe dataset. While all 2 million would be nice, we will ultimately take a sample of recipes anyway to keep our correlation matrix a decent size for this quick analysis. Finally, use the handy `stretch` function from `corrr` to turn the correlation matrix into a long, tidy dataframe where each pair of ingredients has a row.   

```{r recipe-data}
# https://www.kaggle.com/shuyangli94/food-com-recipes-and-user-interactions?select=RAW_recipes.csv
recipes_raw <- read.csv('RAW_recipes.csv', nrows = 10000)

recipes_raw %>% select(name, id, ingredients) %>% head(1)

recipes <- recipes_raw %>%
  select(id, ingredients) %>%
  mutate(ingredients = gsub("\\[|\\]|\\'", '', gsub('"', '', as.character(ingredients))),
         ingredient = sapply(ingredients, function(x) strsplit(x, ', ')[[1]])) %>%
  unnest(cols = ingredient) %>%
  mutate(ingredient = tolower(ingredient)) %>%
  select(-ingredients)

recipes %>% filter(id == 137739)

recipe_ingredient_count <- recipes %>%
  count(ingredient, name = 'recipe_count') %>%
  left_join(chopped_ingredients, by = 'ingredient') %>%
  mutate(frequency = recipe_count/nrow(recipes_raw),
         chopped_ingredient = !is.na(chopped_count))

# 86% of recipes have at least one chopped ingredient
recipes_focus <- recipes %>%
  mutate(chopped_ingredient = ingredient %in% chopped_ingredients$ingredient) %>%
  group_by(id) %>%
  summarise(chopped_flag = any(chopped_ingredient), .groups = 'drop') %>%
  filter(chopped_flag) %>% # keep just recipes including chopped ingredients
  sample_n(size = 1000) # sample down to a reasonable size for computing correlation
  
recipes_wide <- recipes %>%
  filter(id %in% recipes_focus$id) %>% 
  count(id, ingredient) %>%
  pivot_wider(id_cols = 'id', names_from = 'ingredient', values_from = 'n') %>%
  mutate_if(is.numeric, replace_na, 0)

recipes_cor <- recipes_wide %>% 
  select(-id) %>% 
  corrr::correlate(use = 'pairwise.complete.obs', quiet = TRUE)

recipes_cor_long <- corrr::stretch(recipes_cor, na.rm = TRUE) %>%
  left_join(select(recipe_ingredient_count, ingredient, frequency), by = c('x' = 'ingredient')) %>% # add in frequency counts
  left_join(select(recipe_ingredient_count, ingredient, frequency), by = c('y' = 'ingredient')) 

```

From here, we can fill in the dataframe we created earlier containing pairs of Chopped basket ingredients with their correlation value from the Food.com dataset. This represents how often these ingredients occur together in the "real" world - not on Chopped. 

```{r combine}
chopped_ingredients_paired_r <- chopped_ingredients_paired %>%
  left_join(recipes_cor_long, by = c('ingredient' = 'x', 'ingredient2'= 'y')) 

chopped_ingredients_paired_r %>%
  filter(ingredient == 'bok choy' & !is.na(r)) %>%
  head(10) %>%
  arrange(r)
```

The two metrics of most interest are how correlated basket ingredients are with each other, as well as how common they are in the Food.com recipes. Chopped baskets tend to have some more obscure ingredients to make the challenge harder (such as tamarind paste or ladyfingers), so we want to approximate that in our generator as well.  

```{r viz}
chopped_ingredients_paired_r %>%
  filter(!is.na(r)) %>% 
  mutate(round = factor(round, levels = c('appetizer', 'entree', 'dessert'))) %>%
  ggplot(aes(x = round, y = r)) + 
  geom_boxplot() +
  annotate('text', x = 3, y = 0.32, vjust = 1.2, cex = 3,
           label = 'jicama\n+ cilantro') +
  annotate('text', x = 2, y = 0.23, vjust = 1.2, cex = 3,
           label = 'kale\n+ butternut squash') +
  annotate('text', x = 3, y = -0.03, vjust = -0.5, hjust = -0.1, cex = 3,
           label = 'cream cheese\n+ carrots') +
  coord_flip() +
  labs(title = 'Correlation between Chopped ingredient pairs',
       subtitle = 'in Food.com recipes')
```

Ingredients tend to have low correlation values - around 0 or slightly under - but there have been a few lobs in Chopped history (such as the entree round in episode 4 with `country pork ribs, kale, butternut squash, and popcorn on the cob`). Of course, these are just pairwise correlations, so the entire basket isn't taken into account. There could be a pair of nicely correlated ingredients along with two totally out of left field - like in episode 88, where the appetizer round featured `carrots, edamame, buffalo hot dogs, and butterscotch candy`.  

```{r viz-2}
chopped_ingredients_paired_r %>%
  filter(!is.na(frequency.x)) %>%
  select(series_episode, round, ingredient, frequency.x) %>% 
  unique() %>% 
  mutate(round = factor(round, levels = c('appetizer', 'entree', 'dessert'))) %>%
  ggplot(aes(x = round, y = frequency.x)) + 
  geom_boxplot() +
  annotate('text', x = 3, y = 0.108, vjust = 1.3,  cex = 3, label = 'milk') +
  annotate('text', x = 2, y = 0.05, vjust = 1.3,  cex = 3, label = 'honey') +
  annotate('text', x = 1, y = 0.04, vjust = 1.3,  cex = 3, label = 'tomatoes') +
  coord_flip() +
  labs(title = 'Frequency of Chopped ingredients',
       subtitle = 'in Food.com recipes',
       y = '% of Food.com recipes')
```

Chopped ingredients tend to be more rare in the dataset, but the distribution is  right-skewed and long tailed, meaning that most ingredients are "rare." For example, mead is a really low-occuring ingredient from a Chopped basket - but so are sour mix and lima beans. 
In order to guide the generator, let's calculate the average association and frequency in of Chopped ingredients in the Food.com data:  

```{r params}
# so the average similarity and rarity is...
chopped_parameters <- chopped_ingredients_paired_r %>% 
  mutate(round = factor(round, levels = c('appetizer', 'entree', 'dessert'))) %>%
  group_by(round) %>%
  summarise(mean_r = mean(r, na.rm = TRUE),
            sd_r = sd(r, na.rm = TRUE),
            mean_freq = mean(frequency.x, na.rm = TRUE),
            sd_freq = sd(frequency.x, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(lo_r = mean_r - sd_r,
         hi_r = mean_r + sd_r,
         lo_freq = mean_freq - sd_freq,
         hi_freq = mean_freq + sd_freq)

chopped_parameters
```

And then create a function to generate baskets. It's set up to start with a random Chopped ingredient from the selected round (for some Chopped-specific flavor). From there, it looks for ingredients in the Food.com data that match the frequency parameters we set, and that have correlation coefficients within the correlation parameters, and then takes one randomly. Basically, it looks for a rare enough and similar/dissimilar enough ingredient to the one before. Rinse and repeat 3 times and we have a basket of four weirdly paired ingredients!  

```{r function}
generate_chopped_ingredients <- function(input_round) { 

  if (input_round == 'surprise') {
    
    input_round <- sample(c('appetizer', 'entree', 'dessert'), size = 1)
    
  } else if (!input_round %in% c('appetizer', 'entree', 'dessert')) {
    
    return ('Invalid input round.')
    
  }
  
  parameters <- chopped_parameters %>% 
    filter(round == input_round)
  
  first_ingredient <- chopped_ingredients_paired_r %>%
    filter(!is.na(r)) %>%
    count(round, ingredient) %>%
    filter(round == input_round) %>%
    sample_n(size = 1)
  
  generated_ingredients <- c(first_ingredient$ingredient)
  
  for(i in 2:4) {
    
    temp <- recipes_cor_long %>%
      filter(x == generated_ingredients[i-1],
           r > parameters$lo_r[1],
           r < parameters$hi_r[1],
           frequency.y > parameters$lo_freq[1],
           frequency.y < parameters$hi_freq[1]) %>%
      sample_n(size = 1)
    
    generated_ingredients <- c(generated_ingredients, temp$y)
    
  }

  paste0('Open your baskets! Your ', input_round, ' round ingredients are: ', 
           paste(generated_ingredients[1:3], collapse = ', '), ', and ', generated_ingredients[4], '.')
  
}
```

Let's test it out. 

```{r test}
generate_chopped_ingredients(input_round = 'appetizer')
generate_chopped_ingredients(input_round = 'entree')
generate_chopped_ingredients(input_round = 'dessert')
generate_chopped_ingredients(input_round = 'surprise')


```

Good luck chefs.  

--- 

Some notes:  

- Not all Chopped pairs were found in the Food.com data (only 7% were) so it's not a comprehensive picture of all Chopped ingredient pairings.  
- We chose pairwise correlation as the statistic of interest because of the low incidence rate of Chopped ingredients appearing together "in the wild." It would be great to understand how correlated the whole basket is together, but it would be a lot harder to find comparison data/recipes including 3 or all 4 of the ingedients.  
- Because we looked at correlation between pairs of ingredients and not between all four ingredients in the basket, there may be some generated baskets that come out looking pretty easy! The first two ingredients may have low correlation with each other, but then the third may come along and have low correlation with #2 but decently high correlation with #1. 
- The generator also doesn't take into account what _kind_ of ingredient is included. The Chopped producers definitely try to spread out the ingredients to include proteins, sauces, etc. and not just a basket of chicken, beef, shrimp, and pork, or bbq sauce, ketchup, mustard and mayonnaise, for example.  
